---
title: "Homework 5"
author: "Noah Kawasaki"
date: "5/26/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo=FALSE,
  message=FALSE,
  warning=FALSE,
  echo=FALSE,
  root.dir='/Users/noahkawasaki/Desktop/ECON 147/Homework 5'
  )

library(PerformanceAnalytics)
library(quantmod)
library(TSA)
library(car) 
library(tseries) 
library(tidyverse)

symbols <- c("MSFT", "^GSPC")
getSymbols(symbols, from ="2000-01-03", to = "2014-02-21")

# Extract adjusted closing prices
MSFT <- MSFT[, "MSFT.Adjusted", drop=FALSE]
GSPC <- GSPC[, "GSPC.Adjusted", drop=FALSE]

# Calculate log-returns for GARCH analysis
MSFT.ret <- CalculateReturns(MSFT, method="compound")
GSPC.ret <- CalculateReturns(GSPC, method="compound")

# Remove default column header which is treated as the first observation
MSFT.ret <- MSFT.ret[-1,]
GSPC.ret <- GSPC.ret[-1,]

# Rename column
colnames(MSFT.ret) <- "MSFT" 
colnames(GSPC.ret) <- "GSPC"
```


##### (a) Make time plots of the return data from 2000-01-03 to 2014-02-21. Comment on any stylized fact on returns suggested by the plots.
```{r}
# Plot returns
plot(MSFT.ret, main="Microsoft Daily Returns")
plot(GSPC.ret, main="S&P500 Daily Returns")
```

The stylized facts we can notice are:  
- Mean reversion near zero  
- Volatiltiy clustering  


##### (b) For each return series, make a four panel plot containing a return plot, acf, density plot and normal QQ-plot. Do the return series look normally distributed?
```{r}
# Microsoft 
par(mfrow=c(2,2))
plot(MSFT.ret, main="MSFT daily returns")
acf(MSFT.ret, main="ACF", lwd=2)
plot(density(MSFT.ret),type="l", main="Smoothed density", xlab="daily returns", ylab="density estimate", col="slateblue1")
qqnorm(MSFT.ret, col="slateblue1")
qqline(MSFT.ret)
par(mfrow=c(1,1))
```

The ACF of the series indicates no or weak correlation across time. The Density plot shows that it is a bell-shaped curve. However, the QQ-Plot indicates heavier tails than a normal distribution.


```{r}
# S&P500
par(mfrow=c(2,2))
plot(GSPC.ret, main="SP500 daily returns")
acf(GSPC.ret, main="ACF", lwd=2)
plot(density(GSPC.ret),type="l", main="Smoothed density", xlab="daily returns", ylab="density estimate", col="slateblue1")
qqnorm(GSPC.ret, col="slateblue1")
qqline(GSPC.ret)
par(mfrow=c(1,1))
```

The ACF indicates time dependence up to order 2 lages and then some other spikes past that the most likely do not carry economic meaning. The density plot and QQ-Plot also indicate a bell-shaped curve with heavier tails than a normal distribution.

##### (c) Testing normality of each return distribution using Jarque-Bera test statistics.
```{r}
jarque.bera.test(MSFT.ret)
jarque.bera.test(GSPC.ret)
```

For both series, we reject the null hypothesis that r~t~ is normally distributed.


##### (d) Now estimate GARCH(1,1) model parameters (as in Review Questions) and report the estimated values of $\alpha_1$ + $\beta_1$. How do you interpret these results?

#### (e) Plot the fitted values and the observed values. Comment on plots

##### (f) For parameters $\alpha_1$ and $\beta_1$ compute 95% and (asymptotic) confidence intervals.

##### (g) Test H~0~: $\alpha_1$=0 with 95% confidence level for each returns. Do the test H~0~: $\beta_1$=0.9 as well.

### MSFT
```{r}
garch_AP  <- garch(MSFT.ret, order = c(1,1))
n  = dim(MSFT.ret)[1]
n0 = n-1
summary(garch_AP)
attach(garch_AP)

omega_est  = coef[1]                   # estimator of omega
sd_omega   = vcov[1,1]^.5              # standard deviation estimator of the estimator of omega
alpha1_est = coef[2]                   # estimator of alpha1
sd_alpha1  = vcov[2,2]^.5              # standard deviation estimator of the estimator of alpha1
beta1_est  = coef[3]                   # estimator of beta1
sd_beta1   = vcov[3,3]^.5              # standard deviation estimator of the estimator of beta1
sig        = fitted.values[2:n0,c(1)]  # fitted conditioinal standard deviation sigma_t
rsd        = residuals[2:n0]           # fitted residuals for e_t 
fit.r      = sig*rsd                   # fitted cc return for r_t
n1         = n0-1
date       = 1:n1
date_i     = seq(from = 1, to = n0, by = 250)
lab_x      = seq(from = 2000, to = 2014, by = 1)

# plot the real cc return and the fitted cc return
par(mfrow=c(2,1))
par(mar = c(3,4.5,1.5,2))
plot(date, MSFT.ret[3:n], type='l',col='blue',xaxt = 'n', xlab= "",ylab="", xlim=c(1,n1), ylim=c(-.15,.15))
axis(1, at=date_i, labels=lab_x) 
title('Real Data (MSFT)')
plot(date,fit.r, type='l',col='green',xaxt = 'n', xlab= "",ylab="", xlim=c(1,n1), ylim=c(-.15,.15))
axis(1, at=date_i, labels=lab_x) 
title('Fitted Data (MSFT)')
```

For MSFT, $\omega$ = 0.000006, $\alpha_1$ = 0.07, and $\beta_1$ = 0.9. So $\alpha_1$ + $\beta_1$ = 0.97 which is very close to 1 or a random walk process.  

The fitted values and the observed values for both series look very similar, which means the GARCH(1,1) model did a good job at explaining our data.


##### Confidence Intervals
```{r}
a1_ci <- c(alpha1_est-1.96*sd_alpha1, alpha1_est+1.96*sd_alpha1)
a1_ci

b1_ci <- c(beta1_est-1.96*sd_beta1, beta1_est+1.96*sd_beta1)
b1_ci
```


##### Hypothesis Tests
```{r}
# Test a1
(alpha1_est-0)/sd_alpha1

# Test b1
(beta1_est-0.9)/sd_beta1
```

Since both t statistics are greater than the z scores, we reject the null hypotheses and conclude that $\alpha_1$ > 0 and $\beta_1$ $\neq$ 0.9.



### S&P500
```{r}
# GARCH(1,1) with GSPC
garch_AP  <- garch(GSPC.ret, order = c(1,1))
n  = dim(MSFT.ret)[1]
n0 = n-1
summary(garch_AP)
attach(garch_AP)
# The estimation results are in garch_AP
# coef: contains the estimators of omega, alpha1 and beta1
# vcov: contains the variance estimators of the estimators of omega, alpha1 and beta1
# fitted.values: contains the fitted conditioinal standard deviation sigma_t
# residuals: contains the fitted residuals for e_t 

omega_est  = coef[1]                   # estimator of omega
sd_omega   = vcov[1,1]^.5              # standard deviation estimator of the estimator of omega
alpha1_est = coef[2]                   # estimator of alpha1
sd_alpha1  = vcov[2,2]^.5              # standard deviation estimator of the estimator of alpha1
beta1_est  = coef[3]                   # estimator of beta1
sd_beta1   = vcov[3,3]^.5              # standard deviation estimator of the estimator of beta1
sig        = fitted.values[2:n0,c(1)]  # fitted conditioinal standard deviation sigma_t
rsd        = residuals[2:n0]           # fitted residuals for e_t 
fit.r      = sig*rsd                   # fitted cc return for r_t
n1         = n0-1
date       = 1:n1
date_i     = seq(from = 1, to = n0, by = 250)
lab_x      = seq(from = 2000, to = 2014, by = 1)

# plot the real cc return and the fitted cc return
par(mfrow=c(2,1))
par(mar = c(3,4.5,1.5,2))
plot(date, GSPC.ret[3:n], type='l',col='blue',xaxt = 'n', xlab= "",ylab="", xlim=c(1,n1), ylim=c(-.1,.1))
axis(1, at=date_i, labels=lab_x) 
title('Real Data (GSPC)')
plot(date,fit.r, type='l',col='green',xaxt = 'n', xlab= "",ylab="", xlim=c(1,n1), ylim=c(-.1,.1))
axis(1, at=date_i, labels=lab_x) 
title('Fitted Data (GSPC)')
```

For S&P500, $\omega$ = 0.0000015, $\alpha_1$ = 0.086, and $\beta_1$ = 0.9. So $\alpha_1$ + $\beta_1$ = 0.98 which is also very close to 1 or a random walk process.  

The fitted values and the observed values for both series look very similar, which means the GARCH(1,1) model did a good job at explaining our data.


##### Confidence Intervals
```{r}
a1_ci <- c(alpha1_est-1.96*sd_alpha1, alpha1_est+1.96*sd_alpha1)
a1_ci

b1_ci <- c(beta1_est-1.96*sd_beta1, beta1_est+1.96*sd_beta1)
b1_ci
```


##### Hypothesis Tests
```{r}
# Test a1
(alpha1_est-0)/sd_alpha1

# Test b1
(beta1_est-0.9)/sd_beta1
```

$\alpha_1$ is greater than its z score but $\beta_1$ is less than its z score so we reject the null for $\alpha_1$ and conclude that $\alpha_1$ > 0 but fail to reject for $\beta_1$ and conclude that $\beta_1$ = 0.9.


